### 2021-03-15 paper/rebuttal/rebuttal.tex

%
% SECTION-PERFORMANCE COMPARISONS
%
\section{Performance of \textit{word2vec} compared with baselines}
\subsection{Alternatives considered}
In our original manuscript, we compared the performance of what we call the embedding distance---the cosine distance between vectors learned by \textit{word2vec}---to a range of alternatives.
These alternatives included the geographic distance between locations, the distance most commonly used in definitions of the gravity model of mobility, along with the dot product similarity between embedding vectors, and the cosine distance and Jensen-Shannon divergence between vectors obtained from a personalized-page-ranking procedure on the underlying mobility network. Following the advice of the referees, we also add additional baselines to illustrate how the embedding distance performs compared to simpler embedding alternatives and also to a more advanced approach as follows:

\vspace{-0.5cm}
\paragraph{SVD on co-occurrence matrix:}The first of these baselines uses the distance between vectors in an embedding space produced through a singular-value decomposition of the underlying location co-occurrence matrix. First, we construct the co-occurrence matrix of $N$ organizations, $A_{ij}$ given by the co-occurrence of organizations $i$ and $j$ in the same affiliation trajectory.
Then, we apply truncated singular value decomposition with $d=300$ on the flow matrix $A$ directly.

\vspace{-0.3cm}
\paragraph{Laplacian Eigenmap:} Similarly, we consider Laplacian Eigenmap embeddings~\autocite{belkin2003laplacian} as a baseline, which is one of the most fundamental approaches for graph embedding. We generate Laplacian Eigenmap embeddings by applying a truncated singular value decomposition to matrix $L$ given by $L=D - A$, where
$A$ is the co-occurrence matrix of $N$ organizations with element $A_{ij}$ indicating the co-occurrence of organizations $i$ and $j$ in the same affiliation trajectory, and $D$ is the diagonal matrix with diagonals $D_{ii}=\sum_j {A_{ij}}$. We set the dimension to $d=300$.

\vspace{-0.3cm}
\paragraph{Levy's  symmetric SVD \textit{word2vec} distance:} In addition to these suggested baselines, we also consider distances derived from embeddings learned through matrix factorization following the procedure outlined by Levy \& Goldberg~\autocite{levy2014neural}.
Observing that \textit{word2vec} is implicit matrix factorization, Levy proposed a symmetric SVD embedding that directly calculates the embedding that \textit{word2vec} attempts to approximate.
Here, we compare the performance of \textit{word2vec} embeddings against those calculated directly using Levy's SVD.
First, we construct the matrix of $N$ organizations
\begin{equation}
	M_{ij} = \log\left(\frac{N(i,j) |D|}{N(i) N(j)}\right) - \log(k),
\end{equation}
where $N(i,j)$ is the number of times the location pair $(i,j)$ appears given the window size $w$ in the total corpus $D$, $N(i)=\sum_{j=N}^i N(i,j)$ as the number of items $i$ occurred given the window size $w$ in $D$, and $k$ is the number of negative samples. We used $w=1$ and $k=5$ which is the same setting as in our main result. Then, we factorized matrix $M$ with truncated singular value decomposition in the matrix with $d=300$ into $U_d \Sigma_d V_d$, and used the embedding vector as $U_d\sqrt{\Sigma_d}$.

### 2021-03-15 paper/rebuttal/rebuttal.tex

%
% SECTION-PERFORMANCE COMPARISONS
%
\section{Performance of \textit{word2vec} compared with baselines}
\subsection{Alternatives considered}
In our original manuscript, we compared the performance of what we call the embedding distance---the cosine distance between vectors learned by \textit{word2vec}---to a range of alternatives.
These alternatives included the geographic distance between locations, the distance most commonly used in definitions of the gravity model of mobility, along with the dot product similarity between embedding vectors, and the cosine distance and Jensen-Shannon divergence between vectors obtained from a personalized-page-ranking procedure on the underlying mobility network. Following the advice of the referees, we also add additional baselines to illustrate how the embedding distance performs compared to simpler embedding alternatives and also to a more advanced approach as follows:

\vspace{-0.5cm}
\paragraph{SVD on co-occurrence matrix:}The first of these baselines uses the distance between vectors in an embedding space produced through a singular-value decomposition of the underlying location co-occurrence matrix. First, we construct the co-occurrence matrix of $N$ organizations, $A_{ij}$ given by the co-occurrence of organizations $i$ and $j$ in the same affiliation trajectory.
Then, we apply truncated singular value decomposition with $d=300$ on the flow matrix $A$ directly.

\vspace{-0.3cm}
\paragraph{Laplacian Eigenmap:} As the second baseline, we consider Laplacian Eigenmap embeddings~\autocite{belkin2003laplacian}, which is one of the most fundamental approaches for graph embedding. We generate Laplacian Eigenmap embeddings by applying a truncated singular value decomposition to matrix $L$ given by $L=D - A$, where
$A$ is the co-occurrence matrix of $N$ organizations with element $A_{ij}$ indicating the number of the co-occurrences of organizations $i$ and $j$ in the same affiliation trajectory, and $D$ is the diagonal matrix with diagonals $D_{ii}=\sum_j {A_{ij}}$. We set the dimension to $d=300$.

\vspace{-0.3cm}
\paragraph{Levy's  symmetric SVD \textit{word2vec} distance:} In addition to these suggested baselines, we also consider distances derived from embeddings learned through matrix factorization following the procedure outlined by Levy \& Goldberg~\autocite{levy2014neural}.
They utilize SVD to train the {\it word2vec}'s model observing that training \textit{word2vec} can be understood as a factorization of a matrix.
Here, we compare the performance of \textit{word2vec} embeddings against those calculated directly using Levy's SVD.
First, we construct the matrix of $N$ organizations
\begin{equation}
	M_{ij} = \log\left(\frac{N(i,j) |D|}{N(i) N(j)}\right) - \log(k),
\end{equation}
where $N(i,j)$ is the number of times the location pair $(i,j)$ appears given the window size $w$ in the total corpus $D$, $N(i)=\sum_{j=1}^N N(i,j)$ as the number of items $i$ occurred given the window size $w$ in $D$, and $k$ is the number of negative samples. We used $w=1$ and $k=5$ which is the same setting as in our main result. Then, we factorized matrix $M$  into $V_d \Sigma_d U_d$ using truncated SVD with $d=300$, and used $V_d\sqrt{\Sigma_d}$ as the in-vectors.

### 2021-03-15 paper/rebuttal/rebuttal.tex

Furthermore, our subsequent experiments show that our approach's performance is robust, exceeding many other baselines, even direct optimization of the gravity model.
Therefore, we are confident that these changes have brought a qualitative leap in the significance of our findings.
Thus, we would like to request a reconsideration of our manuscript.
Again, we are incredibly appreciative of the referees' comments, which provided the insight and motivation needed to make these breakthroughs.

We have developed the mathematical and theoretical basis of our approach for embedding mobility, and made additional updates to our experiments and manuscript, addressing the three main issues raised by referees:
\vspace{-0.1cm}
\begin{enumerate}
	\itemsep0em
	\item The lack of theoretical justification and grounding for the application of \textit{word2vec}
	\item The lack of simpler baselines to understand the performance of \textit{word2vec}
	\item The lack of sufficient grounding in the literature of neural embeddings and mobility
\end{enumerate}
\vspace{-0.1cm}
We address the first point by presenting the mathematical equivalence between the gravity model of mobility and the skip-gram negative sampling \textit{word2vec} model.
Namely, we demonstrate that \textit{word2vec}'s objective function is equivalent to the gravity model due to the implicit bias of negative sampling,  and thus \textit{word2vec} learns gravity-like relationships between entities, whether it be language or locations.
We believe that this development constitutes a major breakthrough in our understanding of neural embeddings, and in addition to justifying their use in our study, also opens the door for further research applying them to the study of mobility.

For the second point, we have conducted experiments comparing the performance of neural embedding to (\romannum{1}) the singular value decomposition of the mobility network,  (\romannum{2}) a Laplacian eigenmap embedding  (\romannum{3}) a \textit{word2vec} model with matrix factorization,  and  (\romannum{4}) a direct optimization of the gravity model. We find that the embedding distance provided by \textit{word2vec} outperforms all baselines, including the direct optimization of gravity model with the multidimensional scaling, in explaining and predicting real-world mobility.

@@

\subsection{Changes to the manuscript}
To reflect our new findings, we have made the following changes to the manuscript:
\begin{itemize}
	\itemsep0em
	\item Following results demonstrating the performance of the embedding distance compared to geographic distance, we add the section \textbf{word2vec and the gravity model}, which summarizes the link between the two models.
	\item We also added the section \textbf{An implicit bias in the negative sampling} to the methods section of our manuscript.
	\item Since the formulation of the gravity used in the above derivations assumes $\gamma = 1$, we use $\gamma = 1$ for all reported evaluation results. No findings have substantively changed.
	\item We added a sentence to the paper speculating on why cosine similarity works best.
	\item We also added a sentence noting the in-vector and out-vector issue, and speculating on reasons/consequences.
\end{itemize}

%
% SECTION-PERFORMANCE COMPARISONS
%
\section{Performance of \textit{word2vec} compared with baselines}
\subsection{Alternatives considered}
In our original manuscript, we compared the performance of what we call the embedding distance---the cosine distance between vectors learned by \textit{word2vec}---to a range of alternatives.
These alternatives included the geographic distance between locations, the distance most commonly used in definitions of the gravity model of mobility, along with the dot product similarity between embedding vectors, and the cosine distance and Jensen-Shannon divergence between vectors obtained from a personalized-page-ranking procedure on the underlying mobility network. Following the advice of the referees, we also add additional baselines to illustrate how the embedding distance performs compared to simpler embedding alternatives and also to a more advanced approach as follows:

\vspace{-0.5cm}
\paragraph{SVD on co-occurrence matrix:}The first of these baselines uses the distance between vectors in an embedding space produced through the singular value decomposition (SVD) of the underlying location co-occurrence matrix. First, we construct the co-occurrence matrix of $N$ organizations, $A_{ij}$ given by the co-occurrence of organizations $i$ and $j$ in the same affiliation trajectory.
Then, we apply the truncated SVD with $d=300$ on the flow matrix $A$ directly.

\vspace{-0.3cm}
\paragraph{Laplacian Eigenmap:} As the second baseline, we consider Laplacian Eigenmap embeddings~\autocite{belkin2003laplacian}, which is one of the most fundamental approaches for graph embedding. We generate Laplacian Eigenmap embeddings by applying a truncated singular value decomposition to matrix $L$ given by $L=D - A$, where
$A$ is the co-occurrence matrix of $N$ organizations with element $A_{ij}$ indicating the number of co-occurrences of organizations $i$ and $j$ in the same affiliation trajectory, and $D$ is the diagonal matrix with diagonals $D_{ii}=\sum_j {A_{ij}}$. We set the dimension to $d=300$.

\vspace{-0.3cm}
\paragraph{Levy's  symmetric SVD \textit{word2vec} distance:} In addition to these suggested baselines, we also consider distances derived from embeddings learned through matrix factorization following the procedure outlined by Levy \& Goldberg~\autocite{levy2014neural}.
They utilize SVD to train the {\it word2vec}'s model observing that training \textit{word2vec} can be understood as a factorization of a matrix.
Here, we compare the performance of \textit{word2vec} embeddings against those calculated directly using Levy's SVD.
First, we construct the matrix of $N$ organizations
\begin{equation}
	\label{eq:levyword2vec}
	M_{ij} = \log\left(\frac{N(i,j) |D|}{N(i) N(j)}\right) - \log(k),
\end{equation}
where $N(i,j)$ is the number of times the location pair $(i,j)$ appears given the window size $w$ in the total corpus $D$, $N(i)=\sum_{j=1}^N N(i,j)$ as the number of items $i$ occurred given the window size $w$ in $D$, and $k$ is the number of negative samples. We used $w=1$ and $k=5$ which is the same setting as in our main result. Then, we factorized matrix $M$  into $V_d \Sigma_d U_d$ using truncated SVD with $d=300$, and used $V_d\sqrt{\Sigma_d}$ as the in-vectors.

\vspace{-0.3cm}
\paragraph{Direct optimization of gravity model:} Finally, we create an embedding by optimizing for the gravity model directly.
Specifically, we construct a gravity matrix of $N$ organizations, where each cell of the matrix is calculated as $T_{ij}/m_i m_j$, where $T_{ij}$ is the number of co-occurrences of organizations $i$ and $j$ in the same affiliation trajectory, and $m_i$ and $m_j$ are the organizations' populations, defined here as the mean annual number of unique mobile and non-mobile authors affiliated. We then embed the gravity matrix using two approaches: the truncated SVD, and multidimensional scaling (MDS), which embeds each location in an N-dimensional space such that pairwise distances are preserved as well as possible.
Typically, MDS uses Euclidean distance to measure the distance between vectors.

@@

\subsection{New experimental parameters}
In addition to the new baselines, we have also incorporated additional experimental parameters into our analysis.
The first is $\gamma$, a hyper-parameter of the \textit{word2vec} model, which defines the exponent which shapes the negative sampling distribution.
The default of \texttt{gensim} implementation of \textit{word2vec} is set to $\gamma = 0.75$, and this was the value used in our original manuscript.
However, to connect \textit{word2vec} to the gravity model of mobility, we assume $\gamma = 1.0$ (see Section 1).
We now report all main results using $\gamma = 1.0$ and also compare its performance to embeddings learned using $\gamma = 0.75$, as we do with the number of embedded dimensions and window size.
We plot the performance of each combination of hyperparameters by their ability to explain real-world flux of scientific mobility (Fig.~\ref{fig:supp:hyperparameter}), and observe no noticeable difference in performance between values of $\gamma = 0.75$ and $\gamma = 1.0$.
Figures in the updated manuscript have been re-drawn with $\gamma = 1.0$, and the findings remain qualitatively similar.
As a side benefit, including the hyperparameter of $\gamma$ also effectively doubles the number of embeddings trained, and that results are so similar across different embedding provides further evidence of the robustness of the approach applied to our data.

%
% Figure - Hyperparameter performance
%
\begin{figure}[ht!]
	\centering
	\includegraphics[width=0.8\textwidth]{\figdir/Descriptive/hyperparameter_performance.pdf}
	\caption{
		\textbf{Larger dimensions, smaller window size improves embedding performance.}
		The correlation, or amount of flux explained by the embedding distance with varying skip-gram negative sampling hyperparameters.
		Window size refers to $w$, the size of the context window that defines the context in a trajectory.
		Smaller window sizes result in an embedding that explains more flux.
		Embedding dimensions refer to the size of the embedding vector.
		Larger vectors perform better, though with little difference between 200 and 300.
		Gamma refers to the $\gamma$ parameter in \textit{word2vec}, which shapes the negative sampling distribution.
		A value of $\gamma = 0.75$ is the default for \textit{word2vec}.
		There is virtually no difference in performance for different $\gamma$ values.
		All variants perform better on same-country organization pairs, and worse on different-country pairs, than on all pairs of organizations.
		Embeddings with larger dimensions outperform mid-size embeddings for the different-country case.
	}

@@

\subsection{Result}

%
% Figure---Metric performance
%
\begin{figure}[ht!]
	\centering
	\includegraphics[width=0.95\textwidth]{\figdir/Descriptive/distance_metric_performance.pdf}
	\mycaption
	{Neural embeddings outperform baselines for scientific mobility.}
	{
		Cosine distance between embedding vectors generated with \textit{word2vec} explains more of the flux, and better predicts flux when used in the gravity model of mobility than geographic distance and other baselines.
		Shown is the correlation between the flux and embedding distances, measured with $R^{2}$ (left), and the prediction error when using the distance as input to the gravity model of mobility.
		The asterisk denotes the top-performing metric.
		For prediction error, we show results based on both the exponential and power-law forms of the gravity model.
		All embedding-based methods use dimensions of 300.
		Here, embedding distance is obtained from neural embeddings learned with a window size of 1 and $\gamma = 1$.
		In all cases, organization population is defined as the mean annualized number of unique mobile and non-mobile authors, and flux is calculated for all global mobility.
		Baselines include the top-performing distance metrics calculated between vectors obtained by personalized-page rank (PPR), singular value decomposition (SVD), Laplacian eigenmap, direct-factorization following Levy's approach~\autocite{levy2014neural}, and direct optimization of the gravity model using SVD and multidimensional scaling (MDS), as well as the geographic distance between organizations.
		Embedding distance better explains and predicts flux than any other baseline, though there is some variation by experimental parameters (Table~\ref{supp:table:r2_table}, Table~\ref{supp:table:rmse_exp_table}, and Table~\ref{supp:table:rmse_power_table}).
	}
	\label{fig:supp:distancemetrics}
\end{figure}

We observe that the cosine distance between embedding vectors, which we call here the ``Embedding Distance'', explains more of the flux, and results in lower prediction error than all baselines we considered (See Fig.~\ref{fig:supp:distancemetrics}).
This performance advantage holds across \emph{the exponential and the power-law forms of the gravity model of mobility}.
Although the Euclidean distance between directly-factorized vectors performs comparably, and even narrowly beats the embedding distance under certain experimental settings (Tables~\ref{supp:table:r2_table},~\ref{supp:table:rmse_exp_table},~\ref{supp:table:rmse_power_table}); our approach shows the most consistent and robust performance, outperforming even the direct optimization of the gravity models.

\subsection{Why do neural embeddings outperform matrix factorization?}
Why do neural embedding approaches outperform Levy's {\it word2vec} based on matrix factorization?
%Why do neural-embedding approaches, which rely on stochastic gradient descent, outperform direct matrix factorization, especially given that \textit{word2vec} \emph{is} implicitly approximating factorization?
We find that Levy's {\it word2vec} suffers from extreme values in the matrix to factorize, which are numerous when window size is small.
Remind that Levy's {\it word2vec} generates embeddings by factorizing matrix $M$ with entries
$M_{ij} = \log\left[N(i,j)|D|/ (N(i) N(j))\right] - \log k$ based on the SVD (Eq.~\eqref{eq:levyword2vec}), where
$N(i,j)$ is the number of co-occurrences of organizations $i$ and $j$ in the context window.
An issue is that when $N(i,j)$ is closer to zero, $M_{ij}$ goes to $-\infty$.
To prevent such extreme values, a constant is added to $N(i,j)$ if $N(i,j)=0$ in Levy's {\it word2vec}~\autocite{levy2014neural, Qui2018}.
However, this changes the data to which we fit the model, and small $N(i,j)$ values, which are likely to arise due to noise in data, still have a profound impact on the embedding.
The problem is aggravated when the size, $w$, of context window is small.
As the window size becomes smaller, two institutions are less likely to co-appear in the same window, yielding more affiliation pairs with $N(i,j)=0$, which is the case in our experiment (i.e., we set the context window size $w=1$).
%This is particularly problematic when the context window is small and most affiliation pairs would have no direct flow, leading to $N(i,j)=0$ for most affiliation pairs $(i,j)$.
Also, it is well known that SVD itself is vulnerable to outliers \autocite{xu2012robust, huber1981robust, xu1995robust, chandrasekaran2011rank, candes2011robust}.
The stochastic gradient descent algorithm (SGD), which is employed in neural embedding approaches, is more robust than SVD and can enhance the generalization and effectiveness of {\it word2vec}~\autocite{ma2018power, smith2020generalization, zhang2019algorithmic}.

### 2021-03-15 paper/rebuttal/rebuttal.tex

We observe that the cosine distance between embedding vectors, which we call here the ``Embedding Distance'', explains more of the flux, and results in lower prediction error than all baselines we considered (See Fig.~\ref{fig:supp:distancemetrics}).
This performance advantage holds across \emph{the exponential and the power-law forms of the gravity model of mobility}.
Although the Euclidean distance between directly-factorized vectors performs comparably, and even narrowly beats the embedding distance under certain experimental settings (Tables~\ref{supp:table:r2_table},~\ref{supp:table:rmse_exp_table},~\ref{supp:table:rmse_power_table}); our approach shows the most consistent and robust performance, outperforming even the direct optimization of the gravity models.

\subsection{Why do neural embeddings outperform matrix factorization?}
Why do neural embedding approaches outperform Levy's {\it word2vec} based on matrix factorization?
We find that Levy's {\it word2vec} suffers from extreme values in the matrix to factorize, which are numerous when the context window size is small.
Remind that Levy's {\it word2vec} generates embeddings by factorizing matrix $M$ given by Eq.~\eqref{eq:levyword2vec}, i.e.,
\begin{equation}
	M_{ij} = \log\left(\frac{N(i,j) |D|}{N(i) N(j)}\right) - \log k, \nonumber
\end{equation}
based on the SVD where $N(i,j)$ is the number of co-occurrences of organizations $i$ and $j$ in the context window.
An issue is that when $N(i,j)$ is closer to zero, $M_{ij}$ goes to $-\infty$.
To prevent such extreme values, a constant is added to $N(i,j)$ if $N(i,j)=0$ in Levy's {\it word2vec}~\autocite{levy2014neural, Qui2018}.
However, this changes the data to which we fit the model, and small $N(i,j)$ values, which are likely to arise due to noise in data, still have a profound impact on the embedding.
This problem is aggravated when the context window size $w$ is small.
As the window size $w$ becomes smaller, two institutions are less likely to co-appear in the same window, yielding more affiliation pairs with $N(i,j)=0$, which is the case in our experiment (i.e., we set the context window size $w=1$).
Also, it is well known that SVD itself is vulnerable to outliers \autocite{xu2012robust, huber1981robust, xu1995robust, chandrasekaran2011rank, candes2011robust}.
The stochastic gradient descent algorithm (SGD), which is employed in neural embedding approaches, is more robust than SVD and can enhance the generalization and effectiveness of {\it word2vec}~\autocite{ma2018power, smith2020generalization, zhang2019algorithmic}.

### 2021-03-16 paper/rebuttal/rebuttal.tex

\subsection{Changes to the manuscript}
To reflect our new findings, we have made the following changes to the manuscript:
\begin{itemize}
	\itemsep0em
	\item Following results demonstrating the performance of the embedding distance compared to geographic distance, we add the section \textbf{word2vec and the gravity model}, which summarizes the link between the two models.
	\item We also added the section \textbf{An implicit bias in the negative sampling} to the methods section of our manuscript.
	\item Since the formulation of the gravity used in the above derivations assumes $\gamma = 1$, we use $\gamma = 1$ for all reported evaluation results. No findings have substantively changed.
	\item We added a sentence to the paper speculating on why cosine similarity works best.
	\item We also added a sentence noting the in-vector and out-vector issue, and speculating on reasons/consequences.
\end{itemize}

%
% SECTION-PERFORMANCE COMPARISONS
%
\section{Performance of \textit{word2vec} compared with baselines}
\subsection{Alternatives considered}
In our original manuscript, we compared the performance of what we call the embedding distance---the cosine distance between vectors learned by \textit{word2vec}---to a range of alternatives.
These alternatives included the geographic distance between locations, the distance most commonly used in definitions of the gravity model of mobility, along with the dot product similarity between embedding vectors, and the cosine distance and Jensen-Shannon divergence between vectors obtained from a personalized-page-ranking procedure on the underlying mobility network. Following the advice of the referees, we also add additional baselines to illustrate how the embedding distance performs compared to simpler embedding alternatives and also to a more advanced approach as follows:

\vspace{-0.5cm}
\paragraph{SVD on co-occurrence matrix:}The first of these baselines uses the distance between vectors in an embedding space produced through the singular value decomposition (SVD) of the underlying location co-occurrence matrix. First, we construct the co-occurrence matrix of $C$ organizations, $A_{ij}$ given by the co-occurrence of organizations $i$ and $j$ in the same affiliation trajectory.
Then, we apply the truncated SVD with $d=300$ on the flow matrix $A$ directly.

\vspace{-0.3cm}
\paragraph{Laplacian Eigenmap:} As the second baseline, we consider Laplacian Eigenmap embeddings~\autocite{belkin2003laplacian}, which is one of the most fundamental approaches for graph embedding. We generate Laplacian Eigenmap embeddings by applying a truncated singular value decomposition to matrix $L$ given by $L=D - A$, where
$A$ is the co-occurrence matrix of $C$ organizations with element $A_{ij}$ indicating the number of co-occurrences of organizations $i$ and $j$ in the same affiliation trajectory, and $D$ is the diagonal matrix with diagonals $D_{ii}=\sum_j {A_{ij}}$. We set the dimension to $d=300$.

\vspace{-0.3cm}
\paragraph{Levy's  symmetric SVD \textit{word2vec} distance:} In addition to these suggested baselines, we also consider distances derived from embeddings learned through matrix factorization following the procedure outlined by Levy \& Goldberg~\autocite{levy2014neural}.
They utilize SVD to train the {\it word2vec}'s model observing that training \textit{word2vec} can be understood as a factorization of a matrix.
Here, we compare the performance of \textit{word2vec} embeddings against those calculated directly using Levy's SVD.
First, we construct the $C\times C$ matrix $M$ whose entries $M_{ij}$ are given by
\begin{equation}
	\label{eq:levyword2vec}
	M_{ij} = \log\left(\frac{N(i,j) |D|}{N(i) N(j)}\right) - \log(k),
\end{equation}
where $N(i,j)$ is the number of times the location pair $(i,j)$ appears given the window size $w$ in the total corpus $D$, $N(i)=\sum_{j=1}^N N(i,j)$ as the number of items $i$ occurred given the window size $w$ in $D$, and $k$ is the number of negative samples. We used $w=1$ and $k=5$ which is the same setting as in our main result. Then, we factorized matrix $M$  into $V_d \Sigma_d U_d$ using truncated SVD with $d=300$, and used $V_d\sqrt{\Sigma_d}$ as the in-vectors.

\vspace{-0.3cm}
\paragraph{Direct optimization of gravity model:} Finally, we create an embedding by optimizing for the gravity model directly.
Specifically, we construct a gravity matrix of $C$ organizations, where each cell of the matrix is calculated as $T_{ij}/m_i m_j$, where $T_{ij}$ is the number of co-occurrences of organizations $i$ and $j$ in the same affiliation trajectory, and $m_i$ and $m_j$ are the organizations' populations, defined here as the mean annual number of unique mobile and non-mobile authors affiliated. We then embed the gravity matrix using two approaches: the truncated SVD, and multidimensional scaling (MDS), which embeds each location in an N-dimensional space such that pairwise distances are preserved as well as possible.
Typically, MDS uses Euclidean distance to measure the distance between vectors.

### 2021-03-16 paper/rebuttal/rebuttal.tex

\subsection{Changes to the manuscript}
To reflect our new findings, we have made the following changes to the manuscript:
\begin{itemize}
	\itemsep0em
	\item Following results demonstrating the performance of the embedding distance compared to geographic distance, we add the section \textbf{word2vec and the gravity model}, which summarizes the link between the two models.
	\item We also added the section \textbf{An implicit bias in the negative sampling} to the methods section of our manuscript.
	\item Since the formulation of the gravity used in the above derivations assumes $\gamma = 1$, we use $\gamma = 1$ for all reported evaluation results. No findings have substantively changed.
	\item We added a sentence to the paper speculating on why cosine similarity works best.
	\item We also added a sentence noting the in-vector and out-vector issue, and speculating on reasons/consequences.
\end{itemize}

%
% SECTION-PERFORMANCE COMPARISONS
%
\section{Performance of \textit{word2vec} compared with baselines}
\subsection{Alternatives considered}
In our original manuscript, we compared the performance of what we call the embedding distance---the cosine distance between vectors learned by \textit{word2vec}---to a range of alternatives.
These alternatives included the geographic distance between locations, the distance most commonly used in definitions of the gravity model of mobility, along with the dot product similarity between embedding vectors, and the cosine distance and Jensen-Shannon divergence between vectors obtained from a personalized-page-ranking procedure on the underlying mobility network. Following the advice of the referees, we also add additional baselines to illustrate how the embedding distance performs compared to simpler embedding alternatives and also to a more advanced approach as follows:

\vspace{-0.5cm}
\paragraph{SVD on co-occurrence matrix:}The first of these baselines uses the distance between vectors in an embedding space produced through the singular value decomposition (SVD) of the underlying location co-occurrence matrix. First, we construct the co-occurrence matrix of all organizations, $A_{ij}$ given by the co-occurrence of organizations $i$ and $j$ in the same affiliation trajectory.
Then, we apply the truncated SVD with $d=300$ on the flow matrix $A$ directly.

\vspace{-0.3cm}
\paragraph{Laplacian Eigenmap:} As the second baseline, we consider Laplacian Eigenmap embeddings~\autocite{belkin2003laplacian}, which is one of the most fundamental approaches for graph embedding. We generate Laplacian Eigenmap embeddings by applying a truncated singular value decomposition to matrix $L$ given by $L=D - A$, where
$A$ is the co-occurrence matrix of all organizations with element $A_{ij}$ indicating the number of co-occurrences of organizations $i$ and $j$ in the same affiliation trajectory, and $D$ is the diagonal matrix with diagonals $D_{ii}=\sum_j {A_{ij}}$. We set the dimension to $d=300$.

\vspace{-0.3cm}
\paragraph{Levy's  symmetric SVD \textit{word2vec} distance:} In addition to these suggested baselines, we also consider distances derived from embeddings learned through matrix factorization following the procedure outlined by Levy \& Goldberg~\autocite{levy2014neural}.
They utilize SVD to train the {\it word2vec}'s model observing that training \textit{word2vec} can be understood as a factorization of a matrix.
Here, we compare the performance of \textit{word2vec} embeddings against those calculated directly using Levy's SVD.
First, we construct a matrix $M$ whose entries $M_{ij}$ are given by
\begin{equation}
	\label{eq:levyword2vec}
	M_{ij} = \log\left(\frac{N(i,j) |D|}{N(i) N(j)}\right) - \log(k),
\end{equation}
where $N(i,j)$ is the number of times the location pair $(i,j)$ appears given the window size $w$ in the total corpus $D$, $N(i)=\sum_{j=1}^N N(i,j)$ as the number of items $i$ occurred given the window size $w$ in $D$, and $k$ is the number of negative samples. We used $w=1$ and $k=5$ which is the same setting as in our main result. Then, we factorized matrix $M$  into $V_d \Sigma_d U_d$ using truncated SVD with $d=300$, and used $V_d\sqrt{\Sigma_d}$ as the in-vectors.

\vspace{-0.3cm}
\paragraph{Direct optimization of gravity model:} Finally, we create an embedding by optimizing for the gravity model directly.
Specifically, we construct a gravity matrix of all organizations, where each cell of the matrix is calculated as $T_{ij}/m_i m_j$, where $T_{ij}$ is the number of co-occurrences of organizations $i$ and $j$ in the same affiliation trajectory, and $m_i$ and $m_j$ are the organizations' populations, defined here as the mean annual number of unique mobile and non-mobile authors affiliated. We then embed the gravity matrix using two approaches: the truncated SVD, and multidimensional scaling (MDS), which embeds each location in an N-dimensional space such that pairwise distances are preserved as well as possible.
Typically, MDS uses Euclidean distance to measure the distance between vectors.
